RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-09-16 doc/paper.tex
Update

    In this work we present yet another GP--based approach to solve SR, which utilizes the popular iterated local search meta-heuristic~\cite{lourencco2003iterated,lourencco2019iterated}; in the remaining text, we label the metod by \textsc{Rils-Rols}.  The main code of the efficient local search is utilization of the ordinary least square method~\cite{leng2007ordinary} that is in charge for an efficient finding suitable coefficients of linearised formulas.  \fxnote{TODO: maybe expand the core of the algorithm... }

@@

\begin{definition}
  Given is a dataset $D = \{(x_i, y_i)\}_{i=1}^n$, where $\textbf{x} \in \mathbb{R}^d$ represents input variables (features), and $y$ target variable. Suppose that there exists an analytical model of the form $f(\textbf{x})= g^*(\textbf{x}, \theta^*) + \epsilon $ that is a generator of all observations in $D$.  
   The goal of SR is to learn a mapping $\tilde{f}(\textbf{x})=  \tilde{g}(\textbf{x}, \tilde{\theta})  \colon \mathbb{R}^d \mapsto \mathbb{R}$  estimated by searching through the space of (mathematical) expressions  $\tilde{g}$ and parameters $\tilde{\theta}$ where  $\epsilon$ is the presented  white noise using given input data.

@@

Koza~\cite{koza1994genetic} introduced the problem of SR as a specific application of genetic programming. GP deals with the object called programs, which need to be optimized. In particular, the programs are represented by syntax trees consisting of functions/operations over input features and constants; as an example of a function represented by a syntax tree, see Figure~\ref{fig:syntax-tree-example} In essence, syntax trees are elements of the search space of SR. That is, each sample model $\tilde{f}$ may be seen as a point in the search space, represented by respective syntax tree. How accurate is this mapping, we compute in the basis of the historical data $D$ and the chosen error   measure (such as $MSE$, or $R^2$). \fxnote{TODO: proceed...}

@@

   \subsection{Iterated local search framework}

>>>>

    In this work we present yet another GP--based approach to solve SR, which utilizes the popular iterated local search meta-heuristic (ILS)~\cite{lourencco2003iterated,lourencco2019iterated}; in the remaining text, we label the metod by \textsc{Rils-Rols}.  The main code of the efficient local search is utilization of the ordinary least square method (OLS)~\cite{leng2007ordinary} that is in charge for an efficient finding suitable coefficients of linearised formulas.  \fxnote{TODO: maybe expand the core of the algorithm... }

@@

\begin{definition}
  Given is a dataset $D = \{(x_i, y_i)\}_{i=1}^n$, where $\textbf{x} \in \mathbb{R}^d$ represents input variables (features), and $y$ target variable. Suppose that there exists an analytical model of the form $f(\textbf{x})= g^*(\textbf{x}, \theta^*) + \epsilon $ that is a generator of all observations from $D$.  
   The goal of SR is to learn a mapping $\tilde{f}(\textbf{x})=  \tilde{g}(\textbf{x}, \tilde{\theta})  \colon \mathbb{R}^d \mapsto \mathbb{R}$  estimated by searching through the space of (mathematical) expressions  $\tilde{g}$ and parameters $\tilde{\theta}$ where  $\epsilon$ is the presented  white noise using given input data.

@@

Koza~\cite{koza1994genetic} introduced the problem of SR as a specific application of genetic programming. GP deals with the object called programs, which need to be optimized. In particular, the programs are represented by syntax trees consisting of functions/operations over input features and constants; as an example of a function represented by a syntax tree, see Figure~\ref{fig:syntax-tree-example} In essence, syntax trees are elements of the search space of SR. That is, each sample model $\tilde{f}$ may be seen as a point in the search space, represented by respective syntax tree. How accurate is this mapping, we compute in the basis of the historical data $D$ and the chosen error   measure (such as $MSE$, or $R^2$). Interestingly, the search space of SR consists of its discrete and continuous part. More precisely, the model we are seeking for might be obtained as a composition of some functions from a (finite) set of mathematical functions. It is common to use the set of the following elementary mathematical functions in the experiments: $\sqrt{x}, x^2 $, $\sin$, $\cos$, $\log$, $\exp$, $\arcsin$, $\arccos$, and $a^x$. It is allowed to link functions will all four operators: $+$, $-$, $\times$, and $/$. After fixing functions that build models, the set of coefficients that  fits to this model must be provided. For the most physical models, these coefficients may receive any real value, that is they are continuous in its nature.

\fxnote{TODO: proceed...}

@@

   \subsection{Iterated local search}
   \fxnote{TODO: Marko}
